---
title: "Statistical report: The role of CRT as enables for optimizing guideline directed medical therapy (GDMT) in patients with heart failure and reduced ejection fraction (HFrEF): data from the SwedeHF Registry"
subtitle: "Draft"
date: "`r Sys.Date()`"
cover: static/ki_logo_vert_rgb.png
editor: source
format:
  pdf:
    documentclass: scrartcl
    template-partials:
      - "static/before-body.tex"
      - "static/_titlepage.tex"
    include-in-header: 
      - "static/in-header.tex"
    toc: true
    toc-depth: 3
    toc-title: Contents
    lof: true
    lot: true
    cap-location: top
    number-sections: true
    colorlinks: false
    keep-tex: false
bibliography: static/references.bib
csl: static/bmj.csl
link-citations: true
link-bibliography: true
nocite: '@*'
knitr:
  opts_chunk: 
    dev: cairo_pdf
    collapse: true
    comment: "" 
    echo: FALSE
    include: TRUE
    warning: FALSE
    message: FALSE
    fig.pos: "H"
    fig.path: "../output/figs/"
    fig.height: 6
    fig.width: 10
    R.options:
      knitr.graphics.auto_pdf: true
---

```{r}
#| label: set-up-load-data
options(knitr.kable.NA = "")

# load packages, globals and project specific functions
source(here::here("setup/setup.R"))

# load data
load(here("data/clean-data/data.RData"))

# load workbook to write tables to Excel
wb <- loadWorkbook(here("output/tabs/tables.xlsx"))
sheets <- names(wb)

# load pptx file with figs
figs <- officer::read_pptx(path = here::here("output/figs/figs.pptx"))
```           

\newpage

# Data

## Data sources

The following registries were linked using the Swedish personal identity number (PIN)[@pin]: 

- The Swedish Heart Failure Registry (SwedeHF), an ongoing nationwide quality registry started in 2000, 
that includes in- and out-hospital patients with HF, regardless of EF and previous history of HF[@swedehf]. Around 80 variables are recorded at 
discharge from hospital or at the outpatient visit, i.e. index date, and entered into an electronic database managed by the Uppsala Clinical Research 
Center (Uppsala, Sweden). Up to April 2017 the only inclusion criterion was a clinical diagnosis of HF, which was thereafter defined according to the ICD-10 
codes I50.0, I50.1, I50.9, I42.0, I42.6, I42.7, I25.5, I11.0, I13.0, I13.2. In 2021, 69 of 76 Swedish hospitals enrolled patients in SwedeHF, which had a 32% coverage of the prevalent HF population in Sweden[@annualreport]. For this project SHFDB version 4.2.1 was used, see https://kiheartfailure.github.io/shfdb4/. SwedeHF provided information on planned follow-up and the controls. 
- The ICD/PM Registry provided information on date of CRT.
- The National Patient Register (The National Board of Health and Welfare) provided information on comorbidities and the hospitalization outcomes
- The National Prescribed Drug Register (The National Board of Health and Welfare) provided information on the medications
- The Cause of Death Register[@dors] (The National Board of Health and Welfare) provided information for outcomes death
- Longitudinal integrated database for health insurance and labour market studies (LISA)[@lisa] 
and the Register of the Total Population (Statistics Sweden) provided information on the socio-economic factors

## Ethics and Informed consent

The study was approved by the Swedish Ethical Review Authority, dnr 2021-04326 and amendment dnr 2021-06332-02, 2023-05468-02, 2024-02299-02. 
Individual consent was not required, but patients were informed of entry into SwedeHF and able to opt‐out.

\newpage

## Inclusion/exclusion criteria

```{r}
#| label: tbl-flow
#| cache: true
#| cache.comments: false
#| tbl-cap: "Information for flowchart"
#| tbl-pos: "H"

default_kable(flow) %>%
  row_spec(c(1, 11), bold = T) %>%
  column_spec(1, width = "15cm")
```

First patient in: `r min(rsdata$indexdtm)` and last patient in: `r max(rsdata$indexdtm)` from approximately `r rsdata %>% distinct(shf_centre) %>% count() %>% pull(n)` enrolling centers.

Note that a control can receive CRT during the 1 year follow-up and a CRT patient can be explanted (ITT). 

### Limitation 

Patients receiving a CRT can be in the process of being up-titrated (but still not showing up in the last prescription due to the lag). 
Therefore a possible increase in GDMT use can be an effect of this and not an effect of the CRT. 

## Definitions

{{< include src/vars.qmd >}}

\clearpage

### Baseline (index date)

Baseline is defined as date of discharge or, if missing, date of admittance/visit for the 
first encounter in NPR with CRT for the CRT population and registration in SwedeHF for the controls. 
If multiple SwedeHF registrations exists, the registration closest to CRT (within 1 year) is selected. 

### GDMT and loop diuretic

GDMT are: 

- Beta-blocker
- ACEi/ARB/ARNi
- MRA

GDMT and loop diuretic at baseline are defined as a dispensed prescription in NPDR 4 months prior up until day before baseline. 

GDMT and loop diuretic at 1 year follow-up are defined as a dispensed prescription in NPDR 12 months $\pm$ 2 months after baseline. 

The prescription closest to baseline and 1 year is chosen if multiple prescriptions exist. 

The difference between baseline and 1 year follow-up is defined as: 

- Stable: If the patient is not on medication at baseline and not on medication at fu OR stable in achieved % target dose
- Increase: If the patient is not on medication at baseline and on medication at fu OR increase in achieved % target dose. For ACEi/ARB/ARNi a switch from ACEi/ARB to ARNi is also considered an increase. 
- Decrease: If the patient is on medication at baseline and not on medication at fu OR decrease in achieved % target dose. For ACEi/ARB/ARNi a switch from ARNi to ACEi/ARB is also considered an decrease. 

For the combined GDMT variable the sum of the 3 individual medications is calculated (Stable: 0, Increase: 1, Decrease: -1) 
and if the summarized variable is 0 = Stable, >0: Increase, <0: Decrease. 

For the models stable and Decrease where combined into one category forming a dichotomous variable Increase vs. Stable/Decrease for the GDMT and Decrease vs. Stable/Increase for loop diuretics. 

#### Target dose

The dose taken by the patient was evaluated using the strength of the tablet prescribes x the number of pills the patients was prescribed to take each day. The latter is written in free text and an algorithm was derived to extract the dose prescribed. If there was no text or it was not possible to interpret the number of tablets / day was assumed to be the mean number of pills taken per day for the posts with non-missing values with the same ATC code. The % target dose was then calculated according to the defined target dose for each substance (see separate document). For the pills with a combination of substances the target dose was calculated for the MRA, Beta-blocker, ACEi/ARB part of the pill. 

#### Limitations 

1. The text extraction is sometimes a matter of interpretation. 
2. For pills with combination substances the strength was not available. For many combination pills there was only one available dose on the market but for those where there was no information the target dose was set to 50%. The % of combination pills was however low (`r fn(checkcombdoses %>% count(othercomb) %>% mutate(p = n / sum(n) * 100) %>% filter(othercomb) %>% pull(p), 1)`%) except for ARNi (`r fn(checkcombdoses %>% count(arni) %>% mutate(p = n / sum(n) * 100) %>% filter(arni) %>% pull(p), 1)`%). For ARNi, packages containing packages containing 28 or 56 pills were assumed to be of the strength 49/51 together with packages containing 168 pills where the patient was prescribed > 2 pills/day. If the patient was prescribed > 4 pills / day the strength was assumed to be 24/26 mg. The rest was assumed to be strength 97 mg/103 mg. 
3. The last prescription prior to baseline and closes to 1 year was selected, however, the last prescription can be old and the doctor can verbally have instructed the patient to take a different dose/stop medication.

#### Sensitivity analyses

Since the strength of the pills for ARNi were based on an assumption a sensitivity analyses was performed excluding patients recieving ARNi at baseline or fu.

The combined GDMT was also performed were an increase in any of the 3 GDMT was considered an increase, despite the other two descreasing. 

#### Subgroup analyses

Subgroup analyses were performed by systolic blood pressure. 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice] 
and Rubin’s rules was used for combining estimates and standard errors across imputed data sets. 
Variables included in the model are indicated in @tbl-base. CRT was not included in the model but the baseline value of `r paste0(gdmt$label, collapse = ", ")` was. 

## Baseline characteristics

Baseline characteristics are presented by CRT/Controls. 
Categorical variables are presented with n and percentage and tested for 
differences between groups with the chi-square test. 
Continuous variables are presented with median [first quartile-third quartile] and tested for 
differences between groups with the Kruskal-Wallis test. 

## Assoction between CRT and increase in GDMT and decrease in loop diuretic

The association between CRT compared to No CRT (controls) and increase in the individual GDMT and loop diuretic as well as the combined GDMT was evaluated using logistic regression models. 
Partly crude, partly adjusting for the respective baseline medication (for increase in the combined GDMT variable all 3 baseline medications were adjusted for), 
partly adjusting for all variables indicated in @tbl-base together with the baseline medications and 
partly adjusting for all variables together with the change variable for the other GDMT and loop diuretic. 

## Predictors of increase in GDMT and decrease in loop diuretic

The association between baseline characteristics and increase in the individual GDMT as well as the combined GDMT was evaluated using a multivariable logistic regression 
including the variables shown in the forestplots. The variables were selected based on clinical relevance. The difference between CRT and No CRT was evaluated using and interaction effect (the estimates presented are therefore from different multivariable models).  

### Assumptions

Possible outliers were evaluated using Cook's distance and multicollinearity was 
evaluated using the Variance Inflation Factor (VIF) XXX. 

## 1 year follow-up systolic blood pressure and eGFR

For a subset of patients follow-up registrations where available. Follow-up registrations 12 months $\pm$ 2 months 
after baseline were selected where if multiple registrations existed the one closest to 12 months was selected. 
Systolic blood pressure and eGRF was presented descriptively by CRT and the respective medications trajectories for these follow-up registrations. 

# Results

```{r}
med <- rsdata %>%
  summarise(
    med = fn(median(shf_age), dig = 0),
    q1 = fn(quantile(shf_age, probs = 0.25), dig = 0),
    q3 = fn(quantile(shf_age, probs = 0.75), dig = 0)
  ) %>%
  mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
  pull(out)

percfem <- rsdata %>%
  count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 0)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)
```

The median (q1-q3) age is `r med` and `r percfem`% females.    

## Baseline characteristics

{{< include src/base_tab.qmd >}}

\clearpage

## Assoction between CRT and increase in GDMT and decrease in loop diuretic

{{< include src/medbase_tab.qmd >}}

\clearpage

{{< include src/sankey_figs.qmd >}}

\clearpage 

{{< include src/bar.qmd >}}

\clearpage 

{{< include src/med_logreg.qmd >}}

{{< include src/med_logreg_decrease.qmd >}}

{{< include src/med_linjreg.qmd >}}

\clearpage

## Predictors of increase in GDMT and decrease in loop diuretic

{{< include src/predictors.qmd >}}

\clearpage

## 1 year follow-up systolic blood pressure and eGFR

{{< include src/fu.qmd >}}

\clearpage

# Reproducibility

## R code

The R code for all data handling and statistical analyses are found: 

https://github.com/KIHeartFailure/crt-gdmt. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

\clearpage

# References

::: {#refs}
:::
